Citation Analysis of Computer Science

### 2019-11-18 Scientometrics/comp.tex
update

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. The collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have further been postulated as a recent driver for its growth and evolution~\cite{siebel2019_digital}. As also noted in a 2017 US National Academies Report~\cite{nas_2017}, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission".

Given this powerful influence of computer science in research, some understanding its present state, structure, and relationship to other fields that goes beyond satisfying mere interest seems warranted since it would inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy. In using the term inform, we are mindful of (i) Greenhalgh (ii) Sarewitz and Pielke. \emph{I think that I want to cite these two without really knowing how. Greenhalgh for the limitations of positivism. Sarewitz for the limitations of allowing science to do whatever it wants to. I'd also like to see some text about (i) limited funding resources (ii) efficient use of existing resources (iiii) informing policy at multiple levels -- all directed at justifying our study. I'm not sure where to put it though. CS can also be an excellent model system for studying interdisciplinarity cite Porter and colleagues in a study on measuring interdisciplinarity~\cite{porter_measuring_2007} observe that 'The discrepancies between disciplinary roots and research realities loom large', a reference to the challenges associated with knowledge diffusing across researchers who tend to be focused in their own areas. This concern may be especially relevant for computer science given the ubiquity of computing today and limited public funding for research. Maybe squeeze in Tandy's NAS report on interdisciplinarity. I should ping Ohid Yaqub about this too- science policy is his area.}

In a historical study of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles), Salton and Bergmark made the key point that the scientific literature serves a rich source of information to study the structure and historical development of a field~\cite{salton_citation_1979}. They described the global structure of computer science as consisting of three main subareas 
(i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas acknowledged were  (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, (f) non-technical aspects, such as computer education.

@@

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus. the Web of Science research and categorical classifications, the NSF classification system~\cite{nsf_classification}. All three rely on applying one or more journal-derived labels to articles. Disciplinary assignments using journal based classifications have been extensively discussed and critiqued~\cite{wang_large-scale_2016} [need comprehensive list of citations] with Shu and colleagues recently noting from a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Article classification systems have also been developed, however~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, which escape some of the criticisms of journal-based classification.

@@

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only records in DBLP with DOIs (4,291,130) were considered and only matches with a normalized rank of greater than 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. As an example, the following query returns a normalized rank of 0.994.
\begin{lstlisting}
SELECT ts_rank(to_tsvector(`english', `one two three 
four five six seven eight nine ten eleven'), 
plainto_tsquery(`english', `one four three two five 
six seven eight nine ten eleven')); 
\end{lstlisting}

This procedure resulted in a dataset of 634,100 DBLP publications with Scopus identifiers. Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset is referred to as \emph{comp}.\par

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}. 
\textcolor{blue}{Let's provide the exact URL as a citation}
We downloaded v1.2 from the authors' website at `cs.utexas.edu'. 
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus,  We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. [@sitaram please add the number of nodes in the Graclus 20 clusters].

>>>>

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. The collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have further been postulated as a recent driver for its growth and evolution~\cite{siebel2019_digital}. As also noted in a 2017 US National Academies Report~\cite{nas_2017}, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission". Given this powerful influence of computer science in research, some understanding of its present state, structure, and relationship to other fields that seems warranted since it can inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy. For the same reason of its broad influence, computer science can also be an excellent model system for studying interdisciplinarity. These are not novel ideas, however, as far back as 1979, Salton and Bergmark conducted a historical study of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles) in which they made the key point that the scientific literature serves a rich source of information to study the structure and historical development of a field~\cite{salton_citation_1979}. These authors described the global structure of computer science as comprising three main subareas (i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas acknowledged were  (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, (f) non-technical aspects, such as computer education.

@@

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus. the Web of Science research and categorical classifications, the NSF classification system~\cite{nsf_classification}. All three rely on applying one or more journal-derived labels to articles. Disciplinary assignments using journal based classifications have been extensively discussed and critiqued~\cite{wang_large-scale_2016} [need comprehensive list of citations] with Shu and colleagues recently noting from a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Article classification systems have been developed, however~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012} that escape some of the criticisms of journal-based classification.

@@

\emph{misc stuff:  In using the term inform, we are mindful of (i) Greenhalgh (ii) Sarewitz and Pielke. I think that I want to cite these two without really knowing how. Greenhalgh for the limitations of positivism. Sarewitz for the limitations of allowing science to do whatever it wants to. I'd also like to see some text about (i) limited funding resources (ii) efficient use of existing resources (iiii) informing policy at multiple levels -- all directed at justifying our study. I'm not sure where to put it though. CS can also be an excellent model system for studying interdisciplinarity cite Porter and colleagues in a study on measuring interdisciplinarity~\cite{porter_measuring_2007} observe that 'The discrepancies between disciplinary roots and research realities loom large', a reference to the challenges associated with knowledge diffusing across researchers who tend to be focused in their own areas. This concern may be especially relevant for computer science given the ubiquity of computing today and limited public funding for research. Maybe squeeze in Tandy's NAS report on interdisciplinarity. I should ping Ohid Yaqub about this too- science policy is his area.}

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only records in DBLP with DOIs (4,291,130) were considered and only matches with a normalized rank of greater than 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. This procedure resulted in a dataset of 634,100 DBLP publications with Scopus identifiers. Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset is referred to as \emph{comp}.\par

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments.
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus. We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. [@sitaram please add the number of nodes in the Graclus 20 clusters].
